---
title: "Robust coexistence in ecological competitive communities"
author: Pablo Lechon-Alonso, Srilena Kundu, Paula Lemos-Costa, Jose Capitan and Stefano Allesina
output:
  md_document:
    variant: markdown_github
---

## Overview of the code

The folder `code` contains the code to generate random matrices, compute $\alpha_F$ and calculate the transitions to and from feasibility (for small matrices). 

### Required libraries

```{r, warning=FALSE, message=FALSE}
library(sn) # skewed normal distribution
library(mvtnorm) # multinormal distribution
library(tidyverse) # plotting and data wrangling
library(rARPACK) # eigenvalues/vectors of large matrices
library(pracma) # Lambert function
```

### Building random matrices

The file `build_matrices.R` contains the code to build random matrices. To build a matrix, call `build_matrix(n, dist, rho, sk)` with the following parameters:

- `n` size of the matrix
- `dist` either `discrete`, `beta` or `normal`
- `rho` correlation between off-diagonal elements
- `sk` a distribution-dependent parameter controlling skewness

In particular, for `discrete` and `beta`, a parameter `sk == 1` produces a distribution that is centered around its mean, `sk > 1` a right-skewed distribution, and `sk < 1` a left-skewed distribution; for these distributions, only values of `rho` in `1, -1, 0` are implemented.

For the `normal` distribution, `sk = 0` returns the standard normal distribution, and can accept any `rho`; a positive `sk` makes the distribution right-skewed, and a negative `sk` a left-skewed distribution. If `sk` is nonzero, only values of `rho` in `1, -1, 0` are implemented.

All distributions sample coefficients with mean zero and unit variance. Addind a positive constant to the matrices does not impact feasibility.

```{r, warning=FALSE, message=FALSE}
# example 
source("code/build_matrices.R")
set.seed(1)
# example with discrete distribution
B <- build_matrix(1500, "discrete", 0, 1.25)
# exclude diagonal elements to compute stats
Bp <- B
diag(Bp) <- NA
Bp <- as.vector(Bp)
print(mean(Bp, na.rm = TRUE))
print(var(Bp, na.rm = TRUE))

# example with beta distribution
B <- build_matrix(1500, "beta", -1, 1.25)
# exclude diagonal elements to compute stats
Bp <- B
diag(Bp) <- NA
Bp <- as.vector(Bp)
print(mean(Bp, na.rm = TRUE))
print(var(Bp, na.rm = TRUE))

# example with normal distribution
B <- build_matrix(1500, "normal", 0.25, 0)
# exclude diagonal elements to compute stats
Bp <- B
diag(Bp) <- NA
Bp <- as.vector(Bp)
print(mean(Bp, na.rm = TRUE))
print(var(Bp, na.rm = TRUE))
```
## Computing $\alpha_F$

For a given matrix $B$, one can compute the value $\alpha_F$ that would make the equation

$$
(\alpha_F I + B) x = 1
$$
have a positive solution. We provide two options: `get_alpha_F(B, n, rho)` can be used for small matrices $B$, and computes an exact solution by computing the eigenvalues of $n + 1$ matrices; `get_alpha_F_large_random_matrix(B, n, rho)` uses an alternative search algorithm that is better suited for larger matrices.

```{r, warning=FALSE, message=FALSE}
source("code/find_alpha_F.R")
set.seed(1)
# example
B <- build_matrix(25, "normal", 1, 2)
get_alpha_F(B, 25, 1)
get_alpha_F_large_random_matrix(B, 25, 1)
```

### Running simulations

To compute $\alpha_F$ for several random matrices, use `run_simulation`

```{r, eval=FALSE, warning=FALSE, message=FALSE}
set.seed(1)
nsim <- 2500
source("code/build_matrices.R")
source("code/find_alpha_F.R")
source("code/run_simulation.R")
run_simulation(n = 250, dist = "discrete", rho = 1, sk = 1, nsim = nsim)
run_simulation(n = 250, dist = "normal", rho = 0, sk = 0, nsim = nsim)
run_simulation(n = 250, dist = "beta", rho = -1, sk = 1, nsim = nsim)
```

### Computing $p_F$

For each of the three approximations presented in the manuscript (based on small-rank perturbations, using resolvents, using the Richardson's iteration) a file `approximation_small_rank.R` (or `_richardson.R`, or `_resolvent.R`) contains the function `pf_approximations(alpha_values, n, rho, dist, sk)`, which returns the approximate probability of feasibility for the values `alpha_values` (which can be a vector). An example of how to compute these probabilities is contained in `plot_approximations.R`, which computes all three approximations for a simulation of the type specified above.

```{r, warning=FALSE, message=FALSE}
source("code/plot_approximations.R")
plot_approximations("results/n_250_rho_0_dist_normal_sk_0.csv")
```

